---
title: "220C HW1"
author: "Sean Lee"
date: '2022 4 8 '
output: pdf_document
number_sections: true
---
# 1. Cholesky Decomposition
We generate the matrix as follows.
```{r}
# matrix generation

x = seq(0,1,length=400)
dens= function(x,y){
  res = exp(-abs(x-y)/2)
  return(res)
}
R = outer(x,x,FUN=dens)
```

## 1-(a). Inverse matrix
First, we obtain the inverse matrix of $R$ using the 'solve' function in R. Then, we check the required entries.
```{r}
inv.R = solve(R)
inv.R[1,2]; inv.R[2,3]; inv.R[1,3]; inv.R[1,4]
round(inv.R[1:15, 1:15], 3)
```
We can see that this matrix is an example of tridiagonal matrix.

# 1-(b). Cholesky decomposition and eigendecomposition
```{r}
## cholesky decomposition
system.time({
  L = t(chol(R))
  })

## eigendecomposition
system.time(
  {
    E = eigen(R)
  }
)

```
We can see that eigendecomposition requires more time than Cholesky decomposition does.

# 1-(c). Quadratic form computation
```{r}
y = seq(0,2,length=400)

# first: inversion
system.time(
  {
    res1 = t(y)%*%solve(R)%*%y
    res1
  }
)

# second: Cholesky decomposition 
# and backward/forward substitution

system.time(
  {
    res2 = t(y)%*%(backsolve(t(L), forwardsolve(L,y)))
    res2
  }
)

```
We can see that both results are the same, and Cholesky decomposition is faster than the naive inversion.

# 1-(d). Determinant
First, we tried to obtain the $\log|\mathbf{R}|^{-1/2}=-\frac{1}{2}\log|\mathbf{R}|$ based on the direct inversion. Using the determinant function in R, we can directly get the log of the determinant.
```{r}
## first: direct inversion
log.R = determinant(R)$modulus
-.5*log.R
```
We got 1195.048 for the required value. Next, we try to compute based on Cholesky decomposition. Denoting Cholesky decomposition of $\mathbf{R}$ as $\mathbf{R}=\mathbf{L}\mathbf{L}^{T}$ where $\mathbf{L}_{400\times 400}=[L_{ij}]$, we realize that 
$$\log|\mathbf{R}|^{-1/2}=-\frac{1}{2}\sum_{i=1}^{400}\log L_{ii}^2=-\sum_{i=1}^{400}\log L_{ii}.$$
```{r}
## second: Cholesky decomposition
-sum(log(diag(L)))
```
Now, we see a finite value for the result, which implies nonsingularity of $\mathbf{R}$.
